---
output: word_document
---

## UNDERSTANDING DATA AND THEIR ENVIRONMENT: REPORT ASSESSMENT

## Introduction

This report explains the approach taken to forecast sales across stores.
It is reproduble based on code available at ...

The first stage is to load in the data. We will use the tidyverse package and load it with the `library()` function:


```{r, message=FALSE}
library(tidyverse)
# load data
source("code/load-data.R")
```

The task is to estimate sales for the Departments of each store based on the historical training data.

We can check that these files have been loaded into the R envionment with the following command:

```{r}
ls()
```

## Review and describe the data

```{r, out.width="50%"}
b = seq(from = 0, to = 300000, by = 50000)
hist(stores$`Size (sq ft)`, main = "", xlab = "Store size (sqft)", breaks = b)
```

## Joining the data


```{r}
# join store-level data onto training dataset (so we know size)
train_joined = inner_join(train, y = stores)
# would use rename() function to rename columns if needed (not needed)
train_joined = inner_join(train_joined, y = features)
```
## Pre-process the data

## Identify the key factors

## Create the predictive model

There are many packages and approach for forecasting.
We could use the `lm()` function to do a linear regression, for example.
Here we use the xgboost package

```{r}
# install.packages("xgboost")
m1 = lm(Weekly_Sales ~ IsHoliday + Fuel_Price, data = train_joined)
predicted_sales = predict(m1, train_joined)
plot(train_joined$Weekly_Sales, predicted_sales)
```

